
```{r}
# Load Packages
library(Seurat)
library(ggplot2)
library(devtools)
library(RColorBrewer)
library(viridis)
library(circlize)
library(ggpubr)
library(ggrepel)
library(tidyverse)
library(dplyr)
library(phenoclone)
library(nlme)
library(EnhancedVolcano)
library(scales)

# load_all('../../../../TraCR') ## This is just the source code for phenoclone package
```

```{r}
rna_tcr <- readRDS("../../Processed_Data/FROM_STEP_11_Malaria_Specific/rna_tcr_malaria_specific.rds")
```

```{r}
clone_fam_colors <- c(
  "1"="#FDAE61",
  "2"="#FEE08B",
  "3"="#9E0142",
  "4"="#F46D43",
  "5"="#D53E4F",
  "6"="#5E4FA2",
  "7"="#3288BD"
)
```

# 4A Visualize unstim T cells UMAP
```{r, fig.height=3, fig.width=4.5}
# Extract data to make UMAP
umap_data <- FetchData(object = rna_tcr, vars = c("UMAP_1", "UMAP_2", "cell_type","stim","clone_family"))

# Get UMAP
clone_fam_umap_data <- subset(umap_data, (clone_family!="Unassigned") & (stim=="unstimulated") )

clone_fam_umap_plot <- geom_point(data=clone_fam_umap_data, 
                                 aes(x=UMAP_1, y=UMAP_2, group=clone_family, color=clone_family), 
                                 alpha=1, size=0.2)

# Total UMAP
umap_plot <- geom_point(data=umap_data, aes(x=UMAP_1, y=UMAP_2, group=cell_type), 
                        color="lightgray", size=0.01, alpha=0.6)


tiff("../../Plots/Figure_4/4A_clone_fams_unstim_umap.tiff", units="in", width=8, height=6, res=300)

# Plotting
ggplot() + 
  umap_plot + 
  clone_fam_umap_plot + scale_color_manual(values=clone_fam_colors) +
  theme(
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    text = element_text(size = 15)
  )

dev.off()
```

# 4A Visualize iRBC T cells UMAP
```{r, fig.height=3, fig.width=4.5}
# Extract data to make UMAP
umap_data <- FetchData(object = rna_tcr, vars = c("UMAP_1", "UMAP_2", "cell_type","stim","clone_family"))

# Get UMAP
clone_fam_umap_data <- subset(umap_data, (clone_family!="Unassigned") & (stim=="iRBCs") )

clone_fam_umap_plot <- geom_point(data=clone_fam_umap_data, 
                                 aes(x=UMAP_1, y=UMAP_2, group=clone_family, color=clone_family), 
                                 alpha=1, size=0.2)

# Total UMAP
umap_plot <- geom_point(data=umap_data, aes(x=UMAP_1, y=UMAP_2, group=cell_type), 
                        color="lightgray", size=0.01, alpha=0.6)


tiff("../../Plots/Figure_4/4A_clone_fams_irbc_umap.tiff", units="in", width=8, height=6, res=300)

# Plotting
ggplot() + 
  umap_plot + 
  clone_fam_umap_plot + scale_color_manual(values=clone_fam_colors) +
  theme(
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    text = element_text(size = 15)
  )

dev.off()
```

# 4A Visualize bead T cells UMAP
```{r, fig.height=3, fig.width=4.5}
# Extract data to make UMAP
umap_data <- FetchData(object = rna_tcr, vars = c("UMAP_1", "UMAP_2", "cell_type","stim","clone_family"))

# Get UMAP
clone_fam_umap_data <- subset(umap_data, (clone_family!="Unassigned") & (stim=="bead_stimulated") )

clone_fam_umap_plot <- geom_point(data=clone_fam_umap_data, 
                                 aes(x=UMAP_1, y=UMAP_2, group=clone_family, color=clone_family), 
                                 alpha=1, size=0.2)

# Total UMAP
umap_plot <- geom_point(data=umap_data, aes(x=UMAP_1, y=UMAP_2, group=cell_type), 
                        color="lightgray", size=0.01, alpha=0.6)


tiff("../../Plots/Figure_4/4A_clone_fams_bead_umap.tiff", units="in", width=8, height=6, res=300)

# Plotting
ggplot() + 
  umap_plot + 
  clone_fam_umap_plot + scale_color_manual(values=clone_fam_colors) +
  theme(
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    text = element_text(size = 15)
  )

dev.off()
```

# 4B Identify malaria specific T cells
```{r}
exclude_samples <- c("3354 _ sort_Tr1","3410 _ sort_Tr1","3178 _ sort_Tr1", 
                      "3149 _ sort_Tr1", "3178 _ sort_Non", "3410 _ sort_Non", 
                      "3149 _ sort_Non", "3354 _ sort_Non")

unstim_rna_tcr <- subset(rna_tcr, stim == "unstimulated")
unstim_rna_tcr@misc$cell_type_column <- "activation_status"
clonotype_families_unstim <- cluster_clonotypes(unstim_rna_tcr, 
                                                minimum_count = 3, 
                                                exclude_samples = exclude_samples,
                                                n_families = 7,
                                                plot=FALSE)

irbc_rna_tcr <- subset(rna_tcr, stim == "iRBCs")
irbc_rna_tcr@misc$cell_type_column <- "activation_status"
clonotype_families_irbc <- cluster_clonotypes(irbc_rna_tcr, 
                                                minimum_count = 3, 
                                                exclude_samples = exclude_samples,
                                                n_families = 7,
                                                plot=FALSE)

unstim_matrix <- clonotype_families_unstim$matrix
irbc_matrix <- clonotype_families_irbc$matrix

# get clones that are in both stim conditions
unstim_irbc_shared_clones <- intersect(row.names(unstim_matrix), row.names(irbc_matrix))
unstim_matrix <- unstim_matrix[unstim_irbc_shared_clones,]
irbc_matrix <- irbc_matrix[unstim_irbc_shared_clones,colnames(unstim_matrix)]

# Calculate difference
unstim_irbc_diff <- irbc_matrix - unstim_matrix

#### Color clonotypes by family
clonotype_families_df <- rna_tcr@meta.data %>% 
  group_by(cohortid, CTaa, clone_family) %>%
  summarise() %>%
  as.data.frame()

row.names(clonotype_families_df) <- paste(clonotype_families_df$cohortid, clonotype_families_df$CTaa, sep="_")

unstim_irbc_diff <- merge(unstim_irbc_diff, clonotype_families_df, by=0)
####

# Visualize
ggplot(unstim_irbc_diff, aes(x = `TCR-Activated`)) + 
  geom_histogram(bins = 20) + 
  geom_vline(xintercept = 0.4) +
  theme_classic() +
  scale_y_log10()

unstim_v_irbc <- data.frame(
  irbc=irbc_matrix$`TCR-Activated`,
  unstim=unstim_matrix$`TCR-Activated`,
  clone_family=unstim_irbc_diff$`clone_family`
)

tiff("../../Plots/Figure_4/4B_mal_specific_clonotypes.tiff", units="in", width=5.2, height=4, res=300)
ggplot(unstim_v_irbc, aes(x = unstim, y = irbc, color=clone_family)) + 
  geom_jitter(width = 0.025) +
  scale_color_manual(values=c("#FDAE61","#FEE08B","#9E0142","#F46D43","#D53E4F","#5E4FA2","#3288BD","gray") ) +
  geom_abline(intercept = 0.45, slope = 1, color="darkred", size=1) +
  annotate("label",x=0.17, y=0.85, label= "N=39", size=5) + 
  theme_classic() +
  theme(
    axis.title = element_text(size=14),
    axis.text = element_text(size=12)
  ) +
  xlim(-0.026,1) +
  ylim(-0.026,1) +
  xlab("Frequency TCR-Activated\n(Unstimulated)") +
  ylab("Frequency TCR-Activated\n(iRBC-stimulated)")
dev.off()
```

# 4C Pie chart of cell frequencies
```{r}
# Get colors for cell types
cell_to_match_color <- c("Th1","Cytotoxic Th1","Th17","Th2","Tcm",
                         "Naive","Proliferating","TCR-Activated","IFN-Stimulated","Tr1","Treg")

colors <- brewer.pal(11, "Spectral")
names(colors) <- cell_to_match_color
colors["Proliferating"] <- "gray"

# Reorder colors
ordered_cells <- c("Tr1","Treg","Th1","Cytotoxic Th1","Th17","Th2","Tcm","Naive","TCR-Activated","IFN-Stimulated","Proliferating")
colors <- colors[c(ordered_cells)]

# Pie
pie_data_all <- rna_tcr@meta.data %>%
  subset(sample%in%exclude_samples==FALSE) %>%
  subset(stim=="unstimulated") %>%
  group_by(sample, cohortid, cell_type) %>%
  summarise(n=n()) %>%
  group_by(sample, cohortid) %>%
  mutate(total=sum(n))
  
pie_data_all$freq <- pie_data_all$n / pie_data_all$total *100

pie_data_all <- pie_data_all %>%
  as.data.frame() %>%
  complete(cell_type, cohortid) %>%
  group_by(cohortid, cell_type) %>%
  summarise(mean_freq=mean(freq)) %>%
  group_by(cell_type) %>%
  summarise(mean_freq=mean(mean_freq))

# Adjust so sum to 100
pie_data_all$mean_freq <- (pie_data_all$mean_freq / sum(pie_data_all$mean_freq)) * 100

pie_data_all$label <- round(pie_data_all$mean_freq, 2)

# order
pie_data_all <- pie_data_all[match(ordered_cells, pie_data_all$cell_type),]

# Get the positions
pos_df <- pie_data_all %>% 
  mutate(csum = rev(cumsum(rev(mean_freq))), 
         pos = mean_freq/2 + lead(csum, 1),
         pos = if_else(is.na(pos), mean_freq/2, pos))

# Plot
tiff("../../Plots/Figure_4/4C_pie_all.tiff", units="in", width=6, height=6, res=300)

ggplot(pie_data_all, aes(x="", y=mean_freq, fill=fct_inorder(cell_type))) +
  geom_col(width = 1, color = NA) +
  scale_fill_manual(values=unname(colors[ordered_cells])) +
  coord_polar(theta = "y") +
  geom_label_repel(data = pos_df,
                   aes(y = pos, label = paste0(label, "%")),
                   size = 4.5, nudge_x = 1, show.legend = FALSE,
                   segment.color="gray") +
  guides(fill = guide_legend(title = "cell_type")) +
  theme_void() # remove background, grid, numeric labels

dev.off()
```

# 4C Pie chart of malaria specific
```{r}
# Get colors for cell types
cell_to_match_color <- c("Th1","Cytotoxic Th1","Th17","Th2","Tcm",
                         "Naive","Proliferating","TCR-Activated","IFN-Stimulated","Tr1","Treg")

colors <- brewer.pal(11, "Spectral")
names(colors) <- cell_to_match_color
colors["Proliferating"] <- "gray"

# Reorder colors
ordered_cells <- c("Tr1","Th1","Tcm","Cytotoxic Th1","Th17","Th2","Naive","TCR-Activated","IFN-Stimulated","Proliferating","Treg")
colors <- colors[c(ordered_cells)]

# Pie
pie_data_specific <- rna_tcr@meta.data %>%
  subset(sample%in%exclude_samples==FALSE) %>%
  subset(stim=="unstimulated") %>%
  subset(malaria_specific=="yes") %>%
  subset(cell_type!="IFN-Stimulated") %>%
  group_by(sample, cohortid, cell_type) %>%
  summarise(n=n()) %>%
  group_by(sample, cohortid) %>%
  mutate(total=sum(n))
  
pie_data_specific$freq <- pie_data_specific$n / pie_data_specific$total *100

pie_data_specific <- pie_data_specific %>%
  as.data.frame() %>%
  complete(cell_type, cohortid, fill=list(freq = 0)) %>%
  group_by(cohortid, cell_type) %>%
  summarise(mean_freq=mean(freq)) %>%
  group_by(cell_type) %>%
  summarise(mean_freq=mean(mean_freq))

# Adjust so sum to 100
pie_data_specific$mean_freq <- (pie_data_specific$mean_freq / sum(pie_data_specific$mean_freq)) * 100

pie_data_specific$label <- round(pie_data_specific$mean_freq, 2)

# order
pie_data_specific <- pie_data_specific[match(ordered_cells, pie_data_specific$cell_type),]

# drop na
pie_data_specific <- pie_data_specific[is.na(pie_data_specific$mean_freq)==FALSE,]

# Get the positions
pos_df <- pie_data_specific %>% 
  mutate(csum = rev(cumsum(rev(mean_freq))), 
         pos = mean_freq/2 + lead(csum, 1),
         pos = if_else(is.na(pos), mean_freq/2, pos))

# Plot
tiff("../../Plots/Figure_4/4C_pie_malaria_specific.tiff", units="in", width=6, height=6, res=300)

ggplot(pie_data_specific, aes(x="", y=mean_freq, fill=fct_inorder(cell_type))) +
  geom_col(width = 1, color = NA) +
  scale_fill_manual(values=unname(colors[ordered_cells])) +
  coord_polar(theta = "y") +
  geom_label_repel(data = pos_df,
                   aes(y = pos, label = paste0(label, "%")),
                   size = 4.5, nudge_x = 1, show.legend = FALSE,
                   segment.color="gray") +
  guides(fill = guide_legend(title = "cell_type")) +
  theme_void() # remove background, grid, numeric labels

dev.off()
```


# 4D Tr1 Clone activation iRBC vs. resting
```{r}
reorder <- colnames(rna_tcr@assays$RNA@data)
rna_tcr@meta.data <- rna_tcr@meta.data[reorder,]

tr1_clones <- subset(rna_tcr, clone_family=="7")

Idents(tr1_clones) <- "stim"
DefaultAssay(tr1_clones) <- "RNA"
tr1_iRBC_vs_unstim <- FindMarkers(tr1_clones, ident.1 = "iRBCs", ident.2 = "unstimulated", logfc.threshold = 0)

tr1_iRBC_vs_unstim
```

```{r, fig.width=3, fig.height=4}
keyvals <- ifelse(
    tr1_iRBC_vs_unstim$avg_log2FC < -0.5, "#0091ad",
      ifelse(tr1_iRBC_vs_unstim$avg_log2FC > 0.5, "#b7094c",
        'black'))

keyvals[is.na(keyvals)] <- 'black'
names(keyvals)[keyvals == "#0091ad"] <- 'unstimulated'
names(keyvals)[keyvals == "#b7094c"] <- 'iRBCs'
names(keyvals)[keyvals == 'black'] <- 'NS'

tiff("../../Plots/Figure_4/4D_deg_tr1_clones_iRBC_v_unstim.tiff", units="in", width=4.5, height=6.5, res=300)
EnhancedVolcano(tr1_iRBC_vs_unstim,
    lab = rownames(tr1_iRBC_vs_unstim),
    selectLab = c("IFNG","IL10","IL21","IRF4","IL2RA","BCL2A1","NFKB1","RELB","CCL4","CCL3","TNFRSF9","TNFRSF4","NFKBIA","MIR155HG",
                  "CD83","IL21R","MYC","EGR2","MFSD2A",
                  "CD4","CXCR4","CXCR6"),
    x = 'avg_log2FC',
    y = 'p_val_adj',
    colCustom = keyvals,
    FCcutoff = 0.5,
    pCutoff = 0.05,
    labSize = 3.0,
    drawConnectors = TRUE, 
    #xlim = c(-8,8),
    boxedLabels = TRUE
)
dev.off()
```

# 4E Tr1 Clone activation beads vs. resting
```{r}
tr1_beads_vs_unstim <- FindMarkers(tr1_clones, ident.1 = "bead_stimulated", ident.2 = "unstimulated", logfc.threshold = 0)

tr1_beads_vs_unstim
```

```{r, fig.width=3, fig.height=4}
keyvals <- ifelse(
    tr1_beads_vs_unstim$avg_log2FC < -0.5, "#0091ad",
      ifelse(tr1_beads_vs_unstim$avg_log2FC > 0.5, "#e9d8a6",
        'black'))

keyvals[is.na(keyvals)] <- 'black'
names(keyvals)[keyvals == "#0091ad"] <- 'unstimulated'
names(keyvals)[keyvals == "#e9d8a6"] <- 'beads'
names(keyvals)[keyvals == 'black'] <- 'NS'

tiff("../../Plots/Figure_4/4E_deg_tr1_clones_beads_v_unstim.tiff", units="in", width=4.5, height=6.5, res=300)
EnhancedVolcano(tr1_beads_vs_unstim,
    lab = rownames(tr1_beads_vs_unstim),
    selectLab = c("IFNG","IL10","IL21","IRF4","IL2RA","BCL2A1","NFKB1","RELB","CCL4","CCL3","TNFRSF9","TNFRSF4","NFKBIA","MIR155HG",
                  "CD83","IL21R","MYC","EGR2","MFSD2A",
                  "CD4","CXCR4","CXCR6"),
    x = 'avg_log2FC',
    y = 'p_val_adj',
    colCustom = keyvals,
    FCcutoff = 0.5,
    pCutoff = 0.05,
    labSize = 3.0,
    drawConnectors = TRUE, 
    #xlim = c(-8,8),
    boxedLabels = TRUE
)
dev.off()
```


# 4F Tr1 dynamics in 4 infections
```{r}
# Get colors for cell types
cell_to_match_color <- c("Th1","Cytotoxic Th1","Th17","Th2","Tcm",
                         "Naive","Proliferating","TCR-Activated","IFN-Stimulated","Tr1","Treg")

colors <- brewer.pal(11, "Spectral")
names(colors) <- cell_to_match_color
colors["Proliferating"] <- "gray"

# Reorder colors
ordered_cells <- c("Tr1","Treg","Th1","Cytotoxic Th1","Th17","Th2","Tcm","Naive","TCR-Activated","IFN-Stimulated","Proliferating")
ordered_cells <- rev(ordered_cells)
colors <- colors[c(ordered_cells)]

exclude_samples <- c("3354 _ sort_Tr1","3410 _ sort_Tr1","3178 _ sort_Tr1", 
                     "3149 _ sort_Tr1", "3178 _ sort_Non", "3410 _ sort_Non", 
                     "3149 _ sort_Non", "3354 _ sort_Non")

pop_freqs <- rna_tcr@meta.data %>%
  subset(subset=sample%in%exclude_samples==FALSE) %>%
  group_by(sample, cohortid, cell_type, stim) %>%
  summarise(n=n()) %>%
  group_by(sample, cohortid, stim) %>%
  mutate(total=sum(n))

pop_freqs$freq <- pop_freqs$n / pop_freqs$total * 100

pop_freqs_wide <- pivot_wider(pop_freqs, id_cols = c("sample","cohortid","stim"), values_from = "freq", names_from = "cell_type")

pop_freqs_wide[is.na(pop_freqs_wide)] <- 0

pop_freqs_long <- pivot_longer(pop_freqs_wide, names_to = "cell_type", values_to = "freq", cols = unique(pop_freqs$cell_type))

pop_freqs_long$cell_type <- factor(pop_freqs_long$cell_type, levels=ordered_cells)
```

```{r}
sample_meta <- read.csv("../../Metadata/tr1_seq_meta_samples.csv")
days_since <- sample_meta[,c("sample","days_since_first_study_infection")]

# Correct samples in pop_freq
sample_correct <- list(
  "3149_S1"="3149_T1",
  "3149_S2"="3149_T2",
  "3149_S3"="3149_T3",
  "3149_S4"="3149_T4"
)

for ( sample in names(sample_correct) ) {
  
  pop_freqs_long[pop_freqs_long$sample==sample,"sample"] <- sample_correct[sample]

}

pop_freqs_days_study_infect <- merge(pop_freqs_long, days_since, by="sample", all.x=TRUE)
```

```{r, fig.width=3, fig.height=3}
tr1_over_time <- pop_freqs_days_study_infect %>%
  subset(stim=="unstimulated") %>%
  subset(cell_type=="Tr1")
  
# 3481
data_3481 <- tr1_over_time[tr1_over_time$cohortid==3481,] %>%
  mutate(sample = factor(sample, levels = unique(sample)))

p1 <- ggplot(data_3481, aes(x=days_since_first_study_infection, y=freq)) +
  geom_line(color="gray") +
  geom_point(size=3) +
  geom_segment(aes(x = 0 , y = 0, xend = 0, yend = 10), color="darkred", linetype="dashed", size=1) +
  annotate("text", x=0, y=10, label= "mal.", color="darkred", size=5, fontface = 'italic', vjust=-0.5) + 
  theme_classic() +
  theme(
    axis.text = element_text(size = 14, color="black"),
    axis.title = element_text(size = 18, color="black"),
    plot.margin = unit(c(1,0,0,0), 'lines'),
    legend.title=element_text(size=14), 
    legend.text=element_text(size=11)
  ) +
  ylim(0,10) +
  scale_x_continuous(breaks=c(-20,0,20,40,60,80), limits = c(-30,90)) +
  ylab("") +
  xlab("") +
  coord_cartesian(clip = 'off')

# 3410
data_3410 <- tr1_over_time[tr1_over_time$cohortid==3410,] %>%
  mutate(sample = factor(sample, levels = unique(sample)))

p2 <- ggplot(data_3410, aes(x=days_since_first_study_infection, y=freq, group=cohortid)) +
  geom_line(color="gray") +
  geom_point(size=3) +
  geom_segment(aes(x = 0 , y = 0, xend = 0, yend = 10), color="darkred", linetype="dashed", size=1) +
  geom_segment(aes(x = 283 , y = 0, xend = 283, yend = 10), color="darkred", linetype="dashed", size=1) +
  annotate("text", x=0, y=10, label= "mal.", color="darkred", size=5, fontface = 'italic', vjust=-0.5) + 
  annotate("text", x=283, y=10, label= "mal.", color="darkred", size=5, fontface = 'italic', vjust=-0.5) + 
  theme_classic() +
  theme(
    axis.text = element_text(size = 14, color="black"),
    axis.title = element_text(size = 18, color="black"),
    axis.text.x=element_text(),
    plot.margin = unit(c(1,0,0,0), 'lines'),
    legend.title=element_text(size=14), 
    legend.text=element_text(size=11)
  ) +
  ylim(0,10) +
  scale_x_continuous(breaks=c(0,100,200,300), limits = c(-50,350)) +
  ylab("") +
  xlab("") +
  coord_cartesian(clip = 'off')

# 3178
data_3178 <- tr1_over_time[tr1_over_time$cohortid==3178,] %>%
  mutate(sample = factor(sample, levels = unique(sample)))

p3 <- ggplot(data_3178, aes(x=days_since_first_study_infection, y=freq, group=cohortid)) +
  geom_line(color="gray") +
  geom_point(size=3) +
  geom_segment(aes(x = 0 , y = 0, xend = 0, yend = 10), color="darkred", linetype="dashed", size=1) +
  geom_segment(aes(x = 379 , y = 0, xend = 379, yend = 10), color="darkred", linetype="dashed", size=1) +
  annotate("text", x=0, y=10, label= "mal.", color="darkred", size=5, fontface = 'italic', vjust=-0.5) + 
  annotate("text", x=379, y=10, label= "mal.", color="darkred", size=5, fontface = 'italic', vjust=-0.5) + 
  theme_classic() +
  theme(
    axis.text = element_text(size = 14, color="black"),
    axis.title = element_text(size = 18, color="black"),
    axis.text.x=element_text(),
    plot.margin = unit(c(0,0,0,0), 'lines'),
    legend.title=element_text(size=14), 
    legend.text=element_text(size=11)
  ) +
  ylim(0,10) +
  scale_x_continuous(breaks=c(0,200,400), limits = c(-50,450)) +
  ylab("") +
  xlab("") +
  coord_cartesian(clip = 'off')

# 3354
data_3354 <- tr1_over_time[tr1_over_time$cohortid==3354,] %>%
  mutate(sample = factor(sample, levels = unique(sample)))

p4 <- ggplot(data_3354, aes(x=days_since_first_study_infection, y=freq, group=cohortid)) +
  geom_line(color="gray") +
  geom_point(size=3) +
  geom_segment(aes(x = 0 , y = 0, xend = 0, yend = 10), color="darkred", linetype="dashed", size=1) +
  geom_segment(aes(x = 61 , y = 0, xend = 61, yend = 10), color="darkred", linetype="dashed", size=1) +
  geom_segment(aes(x = 148 , y = 0, xend = 148, yend = 10), color="darkred", linetype="dashed", size=1) +
  annotate("text", x=0, y=10, label= "mal.", color="darkred", size=5, fontface = 'italic', vjust=-0.5) + 
  annotate("text", x=61, y=10, label= "mal.", color="darkred", size=5, fontface = 'italic', vjust=-0.5) + 
  annotate("text", x=148, y=10, label= "mal.", color="darkred", size=5, fontface = 'italic', vjust=-0.5) + 
  theme_classic() +
  theme(
    axis.text = element_text(size = 14, color="black"),
    axis.title = element_text(size = 18, color="black"),
    axis.text.x=element_text(),
    plot.margin = unit(c(0,0,0,0), 'lines'),
    legend.title=element_text(size=14), 
    legend.text=element_text(size=11)
  ) +
  ylim(0,10) +
  scale_x_continuous(breaks=c(-75,0,75,150), limits = c(-80,180)) +
  ylab("") +
  xlab("") +
  coord_cartesian(clip = 'off')

# plot
fig <- ggarrange(p1, p2, p3, p4,
          ncol = 2, nrow = 2,
          align = "hv")

tiff("../../Plots/Figure_4/4F_tr1_dynamics.tiff", units="in", width=6, height=6, res=300)

annotate_figure(fig, left = text_grob("% Tr1 of CD45RA-", rot = 90, vjust = 1.5, hjust=0.4, size=20),
                    bottom = text_grob("Days", vjust = -0.8, hjust=0, size=20))

dev.off()
```


# 4G frequency change following infection
```{r}
exclude_samples <- c("3354 _ sort_Tr1","3410 _ sort_Tr1","3178 _ sort_Tr1", 
                     "3149 _ sort_Tr1", "3178 _ sort_Non", "3410 _ sort_Non", 
                     "3149 _ sort_Non", "3354 _ sort_Non")

pop_freqs <- rna_tcr@meta.data %>%
  subset(subset=sample%in%exclude_samples==FALSE) %>%
  group_by(sample, cohortid, cell_type, stim) %>%
  summarise(n=n()) %>%
  group_by(sample, cohortid, stim) %>%
  mutate(total=sum(n))

pop_freqs$freq <- pop_freqs$n / pop_freqs$total * 100

pop_freqs_wide <- pivot_wider(pop_freqs, id_cols = c("sample","cohortid","stim"), values_from = "freq", names_from = "cell_type")

pop_freqs_wide[is.na(pop_freqs_wide)] <- 0

pop_freqs_long <- pivot_longer(pop_freqs_wide, names_to = "cell_type", values_to = "freq", cols = unique(pop_freqs$cell_type))

pop_freqs_long$cell_type <- factor(pop_freqs_long$cell_type, levels=ordered_cells)
```

```{r}
sample_meta <- read.csv("../../Metadata/tr1_seq_meta_samples.csv")
days_since <- sample_meta[,c("sample","days_since_malaria_at_least")]

pop_freqs_days_since <- merge(pop_freqs_long, days_since, by="sample", all.x=TRUE)
```


```{r, fig.width=3, fig.height=3}
tr1_over_time <- pop_freqs_days_since %>%
  subset(stim=="unstimulated") %>%
  subset(days_since_malaria_at_least!=0) %>%
  subset(cell_type=="Tr1")
  
tr1_over_time_lme <- tr1_over_time %>%
  subset(cohortid%in%c(3125,3158)==FALSE)

lme1 <- lme(freq ~ log(days_since_malaria_at_least)*cohortid, data=tr1_over_time_lme, random= ~1|cohortid)

summary(lme1)

tr1_over_time$cohortid <- factor(tr1_over_time$cohortid, 
                                 levels = c(3125, 3158, 3377, 
                                            3354, 3481, 3178, 3410, 3528, 3149))

tiff("../../Plots/Figure_4/4G_tr1_over_time.tiff", units="in", width=6, height=6, res=300)

ggplot(data=tr1_over_time, aes(x=days_since_malaria_at_least, y=freq, color=cohortid)) +
  geom_point(size=3) +
  geom_smooth(method="glm", formula= (y ~ log(x)), se=FALSE, size=1.5 )+
  #geom_function(fun = ~ -2.7*log(.x) +15.42 ) +
  #scale_x_continuous(trans = "log10") +
  scale_color_manual(values=c("#b5e48c","#52b69a","#168aad",
                              "#ffba08","#f48c06","#e85d04","#d00000","#9d0208","#6a040f")) +
  theme_classic() + 
  theme(
    #legend.position="none",
    axis.text = element_text(size = 17, color="black"),
    axis.title.x = element_text(size = 20, color="black", vjust = 0),
    axis.title.y = element_text(size = 20, color="black", vjust = 0),
    legend.position=c(.82,.78),
    legend.title=element_blank(),
    legend.text=element_text(size=16)
  ) +
  xlab("Days Since Malaria") +
  ylab("% Tr1 of CD45RA-")

dev.off()

```

# 4H correlation between Tr1 freq and parasitemia
```{r}
expansion_qpcr <- read.csv("../../Metadata/correlating_qpcr_and_tr1_expansion.csv")

expansion_qpcr$diff <- expansion_qpcr$tr1_after - expansion_qpcr$tr1_before
expansion_qpcr$log_qpcr <- log10(expansion_qpcr$qPCR)

expansion_qpcr$fc <- expansion_qpcr$tr1_after / expansion_qpcr$tr1_before

reg_plot <- ggplot(data=expansion_qpcr, aes(x=log_qpcr, y=diff)) +
  geom_point(size=3) + 
  stat_smooth(method='lm', formula= y~x) +
  stat_cor(method = "pearson", size=6) +
  theme_classic() + 
  theme(
    #legend.position="none",
    axis.text = element_text(size = 16, color="black"),
    axis.title.x = element_text(size = 20, color="black", vjust = 0),
    axis.title.y = element_text(size = 20, color="black", vjust = 0),
    legend.title=element_blank(),
    legend.text=element_text(size=15)
  ) +
  xlab("Parasitemia, Log(qPCR)") +
  ylab("Change in % Tr1")

tiff("../../Plots/Figure_4/4H_parasitemia_corr.tiff", units="in", width=6, height=6, res=300)
reg_plot
dev.off()

# Temperature not as good a predictor as 
model <- lm(formula = diff ~ log_qpcr + temperature, data=expansion_qpcr)

summary(model)
```









